---
output: html_document
---
# Lecture 7
1. Recap:
  * The base: plot() functions
2. Today's agenda
  * ggplot2()

Three main modes for plotting in R:

  - plot functions
```{r}
x <- 0:64/64
y <- sin(3*pi*x)
plot(x, y, type = "l", col = "blue",
    main = "points with bg & legend(*, pt.bg)")
points(x, y, pch = 21, bg = "white")
legend(.4,1, "sin(c x)", pch = 21, pt.bg = "white", lty = 1, col = "blue")
```
  - lattice functions

  - ggplot functions

# ggplot2 package
What is ggplot2?

  1. Its a package in R called ‘ggplot2’

  2. An implementation of the grammar of graphics: a description of how a graphics can be broken down into abstract concepts.
  
  From ggplot2 book:
  "In brief, the grammar tells us that a statistical graphic is a mapping from data to aesthetic attributes (colour, shape, size) of geometric objects (points, lines, bars). The plot may also contain statistical transformations of the data and is drawn on a specific coordinate system."

  3. Written by Hadley Wickham

### qplot()
  - Stands for quick plot
  - Looks very similar to plot() function
  - Needs a data frame.
  - Plots are made up of aesthetics(size, shape, color) and geoms(points, lines). 
  - Geoms: geometric objects to represent observations
  - Factors are very important in ggplot: make sure they are labeled in a meaningful way. 
  - ggplot is the core function and can do things that qplot() can not do.

#### An example
The diamond example: the data set is available under the package ggplot2.

This dataset containing the prices and other attributes of almost 54,000 diamonds.
```{r}
library(ggplot2)
qplot(carat, price, data = diamonds)
# compared to
plot(diamonds$carat,diamonds$price, type = 'p')
qplot(log(carat), log(price), data = diamonds)
dsmall <- diamonds[sample(nrow(diamonds), 300), ]
qplot(log(carat),log(price), colour = color, data = dsmall)
```

```{r}
#diamonds example
qplot(carat, price, data = diamonds)
qplot(log(carat), log(price), data = diamonds)
qplot(carat, x * y * z, data = diamonds)

set.seed(1410)
dsmall <- diamonds[sample(nrow(diamonds), 100), ]
qplot(carat, price, data = dsmall)
qplot(log(carat), log(price), data = dsmall)

#modifying aesthetics: separate groups by the color of the diamonds. 
qplot(log(carat),log(price), colour = color, data = dsmall)
# separate groups by the shape
qplot(log(carat),log(price), shape = cut,data = dsmall)


#geom arguments adding some statistics "smoothing the data"
qplot(carat, price, geom = c("point", "smooth"), data = dsmall)

#histogram
qplot(carat, data = diamonds, geom = "histogram")
qplot(carat, data = diamonds, geom = "density")

#histogram
qplot(carat, data = diamonds, geom = "density", colour = color)
qplot(carat, data = diamonds, geom = "histogram", fill = color)

#facets
qplot(carat, data = diamonds, geom = "density",facets = color~.)
qplot(carat, data = diamonds, geom = "histogram", colour = color, facets = color~.)
qplot(carat,price, data = diamonds, facets = color~.)
qplot(carat,price, data = diamonds, facets = .~cut)
qplot(log(carat), log(price), data = dsmall, facets = cut~., 
      geom = c("point", "smooth"), method = 'lm', colour = cut)
```

### ggplot()
When we used qplot(), it did a lot of things for us: it created a plot object, added layers, and displayed the result, using many default values along the way.

#### Some basics about ggplot() function:
  1. Needs a data frame
  2. aesthetic mappings: size, color
  3. geoms: geometric objects
  4. facets: multiple panels
  5. stats: statistical transformations

#### Plots are built up in layers:
  1. Plot the data
  2. Overlay a summary: a statistics summary, regression line
  3. Annotations
  
We just learned that:
```{r}
qplot(log(carat), log(price), data = dsmall, facets = cut~., 
      geom = c("point", "smooth"), method = 'lm', colour = cut)
```
Now, we learn how to build with ggplot() function:
Building ggplot(), adding things "piece by piece": 
```{r}
g = ggplot(dsmall, aes(log(carat), log(price))) 
g = ggplot(dsmall, aes(log(carat), log(price))) +  geom_point()
```

Adding a smoother:
```{r}
g + geom_point() + geom_smooth()
```
Adding another smoother:
```{r}
g + geom_point() + geom_smooth(method = 'lm')
```
Adding Facets:
```{r}
g + geom_point() +facet_grid(cut~.)+geom_smooth(method = 'lm')
```
Adding Colors:
```{r}
g + geom_point(aes(colour = cut)) +facet_grid(cut~.) + geom_smooth(method = 'lm')
```

```{r}
g1 <- ggplot(dsmall, aes(log(carat), log(price), colour = cut))
g1 + geom_point() +facet_grid(cut~.) + geom_smooth(method ='lm')
```
### Remarks:

  1. Orders do not matter
  
  2. Make sure that your data has meaningful factor levels.

### Some other changes could be done to your plot:
  1. Modifying Aesthetics:
  ```
  g + geom_point(aes(colour = cut), size = 4, alpha = 1/4)
  ```
  2. Annotation:
  You can change the labels of x, y axis by adding xlab(), ylab(), ggtitle()
  1. The background of your ggplot can be changed using
      ```
      theme_gray() default setting
      theme_bw() ---change to black
      ```
      ```{r}
      g  + geom_point(aes(colour = cut)) + facet_grid(cut~.) + 
      geom_smooth(method = 'lm', aes(colour = cut)) + theme_gray()
      ```
  2. Change the font by  `{r} base_family = "Times"`
  3. You can also delete your legend by issuing
      ```
      theme(lengend.position = "none"")
      ```
